Basic regression in Gaussian processes

### 2019-11-01 notebook/complete-discrete-version.Rmd
rename and clean up

damage <- 0.05
control <- 1
discount <- 0.999

n_s <- 121
states <- seq(0, 120, length = n_s)
actions <- seq(0, 120, length = n_s)

@@

may <- function(a){  
  function(x, h = 0){ # May
    y <- x - h
    pmax(
      ## controlling h is controlling the bifurcation point directly...
      y + y * p$r * (1 - y / p$K)  - a * y ^ p$q / (y ^ p$q + p$b ^ p$q),  
      0)
  }
}

reward_fn <- function(x,h) - damage * (x / 100) ^ 2 - control * (h / 100)

@@

true_a <- 27.2      ### 21.5
believe_a <- 29 # 28.5   ### 22.5

@@

# Learning

```{r}
x0 <- which.min(abs(states - 25))
```

@@

```{r}
set.seed(12345)
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = 100)
df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

@@

```{r}
set.seed(12345)
df <- mdp_planning(m_belief, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = 1000)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

@@

```{r}
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = true_p$policy, x0 = x0, Tmax = 1000)

df %>% 
  mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + 
  geom_point(alpha=0.5)  + 
  geom_line(aes(time, action), col="blue")
```

@@

```{r}
wd <- sd(possible_a)
prior <- dnorm(possible_a, believe_a, wd/2)
prior <- prior / sum(prior)
```

@@

```{r}
mdp_learning_ <- memoise::memoise(mdp_learning)
  learning_sim <- mdp_learning_(transition, reward, discount, 
                      x0 = x0, 
                      Tmax = Tmax, 
                      true_transition = transition[[true_i]],
                      model_prior = prior,
                      type = "value iteration", 
                      epsilon = 1e-2)

@@

```{r}
 learning_sim$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 5)) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = as.numeric(param)) %>% 
  ggplot(aes(param, probability, group = time, alpha = time)) + 
  geom_line()
```

>>>>

damage <- .5
control <- 1
endemic <- 50
discount <- 0.999 # Note that dynamics are not on scale of years but ~ days
e <- .4
n_s <- 121
states <- seq(0, 120, length = n_s)
actions <- seq(0, 120, length = n_s)

@@

may <- function(a){  
  function(x, h = 0){ # May
    y <- x - e * h
    pmax(
      ## controlling h is controlling the bifurcation point directly...
      y + y * p$r * (1 - y / p$K)  - a * y ^ p$q / (y ^ p$q + p$b ^ p$q),  
      0)
  }
}

## safe state
##

reward_fn <- function(x,h) - (damage * pmax(x - endemic, 0)) ^ 2  - (control * h)
# reward_fn <- function(x,h) - damage * (x / 100) ^ 2 - control * (h / 100) ^ 2
#

@@

true_a <- 27.4      ### 21.5
believe_a <- 28.5 # 28.5   ### 22.5

@@

# Planning

```{r}
x0 <- which.min(abs(states - 5))
Tmax <- 700
```

@@

```{r}
set.seed(12345)
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = Tmax)
```

```{r}
df %>% 
  mutate(state = states[state], action = actions[action] * 5) %>% 
  select(-obs,-value) %>%
  pivot_longer(-time) %>%
  ggplot(aes(time, value, col=name)) + 
  geom_point() + geom_path() #+ 
  #facet_wrap(~name, scales="free_y", ncol=1)

@@

```{r}
set.seed(12345)
df <- mdp_planning(m_belief, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = Tmax)

df %>% 
  mutate(state = states[state], action = actions[action] * 5) %>% 
  select(-obs,-value) %>%
  pivot_longer(-time) %>%
  ggplot(aes(time, value, col=name)) + 
  geom_point() + geom_path()
```

@@

```{r}
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = true_p$policy, x0 = x0, Tmax = Tmax)

df %>% 
  mutate(state = states[state], action = actions[action] * 5) %>% 
  select(-obs,-value) %>%
  pivot_longer(-time) %>%
  ggplot(aes(time, value, col=name)) + 
  geom_point() + geom_path()
```

@@

```{r}
wd <- sd(possible_a)
prior <- dnorm(possible_a, believe_a, wd/3)
prior <- prior / sum(prior)
```

@@

```{r}
mdp_learning_ <- memoise::memoise(mdp_learning)
  learning_sim <- mdp_learning_(transition, reward, discount, 
                      x0 = x0, 
                      Tmax = 20, 
                      true_transition = transition[[true_i]],
                      model_prior = prior,
                      type = "value iteration", 
                      epsilon = 1e-2)

@@

```{r}
 learning_sim$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 5)) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = possible_a[as.integer(param)]) %>% 
  ggplot(aes(param, probability, group = time, alpha = time)) + 
  geom_line()
```

### 2019-11-01 notebook/gp-manual-vs-greta.Rmd
rename and clean up

```{r}
cov <- kernel
sigma.n <- posterior_means["obs_sd"]
cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% greta_array(obs$y)
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
```

>>>>

```{r}
cov <- kernel
sigma.n <- posterior_means["obs_sd"]

cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% greta_array(obs$y)
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)

Ef <- calculate(Ef, 
                values = list(rbf_len = posterior_means["rbf_len"], 
                              rbf_var = posterior_means["rbf_var"])) %>% as.numeric()
Cf <- calculate(Cf, 
                values = list(rbf_len = posterior_means["rbf_len"], 
                              rbf_var = posterior_means["rbf_var"]))
```

```{r}
## Theory/Analytic
gp <- data.frame(x_predict = x_predict, Ef = Ef, sigma = 2*sqrt(diag(Cf)) )
## Draws
values <- mvrnorm(200, Ef, Cf)
dat <- data.frame(x=x_predict, t(values)) %>%  pivot_longer(-x)
```

```{r}
ggplot(dat,aes(x=x,y=value)) +
  geom_ribbon(data=gp, 
              aes(x=x_predict, y=Ef, ymin=Ef-sigma, ymax=Ef+sigma),
              fill="grey80") + # Var
  geom_line(aes(group=name), alpha=0.3, col=rgb(0.7, 0.1, 0.4, 0.1)) + #REPLICATES
  geom_line(data=gp,aes(x=x_predict,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
```

### 2019-11-01 notebook/greta-gp.Rmd
rename and clean up

```{r}
# kernel & GP
y <- diff(ex$x)
x <- ex$x[-n]
ggplot(data.frame(x,y), aes(x,y)) + geom_point()

# hyperparameters
rbf_var <- lognormal(0, 1)
rbf_len <- lognormal(0, 10)
obs_sd <- lognormal(0, .1)

@@

# prediction
f_plot <- project(f, x_predict)
```

```{r}
# fit the model by Hamiltonian Monte Carlo
m <- model(f, rbf_len, rbf_var, obs_sd)
draws <- mcmc(m, chains = 1, one_by_one = TRUE)
```

```{r}
df <- draws[[1]] %>% as_tibble(.name_repair = "universal") 
hyperpars <- df %>% dplyr::select(rbf_len, obs_sd, rbf_var)  
posterior <- df %>% dplyr::select(-rbf_len, -obs_sd, -rbf_var)

hyperpars %>% colMeans()
#hyperpars %>% gather() %>% ggplot(aes(x = value, fill=key)) + stat_density() + facet_wrap(~key)
```

```{r}
posterior %>% t() %>% as_tibble() %>% dplyr::select(1:200) %>% mutate(x = x_predict) %>% pivot_longer(-x)  %>%
  ggplot(aes(x,value)) + 
  geom_line(aes(group=name), alpha=0.3, col=rgb(0.7, 0.1, 0.4, 0.1)) +
  geom_point(data = obs, aes(x,y))
```

```{r}
```

>>>>

```{r}
# kernel & GP
y <- diff(ex$x)
x <- ex$x[-n]
obs <- data.frame(x,y)
ggplot(obs, aes(x,y)) + geom_point()
```

```{r}
# hyperparameters
rbf_var <- lognormal(0, 1)
rbf_len <- lognormal(0, 10)
obs_sd <- lognormal(0, .1)

@@

# prediction
x_predict <- seq(0, 150, len=200)
f_plot <- project(f, x_predict)
```

```{r}
# fit the model by Hamiltonian Monte Carlo over hyperparameters only, not sampling from the GP
m <- model(rbf_len, rbf_var, obs_sd)
draws <- mcmc(m, chains = 1, one_by_one = TRUE)
```

```{r}
df <- draws[[1]] %>% as_tibble(.name_repair = "universal") 
readr::write_csv(df, "draws.csv.gz")
hyperpars <- df %>% dplyr::select(rbf_len, obs_sd, rbf_var)
posterior_means <- hyperpars %>% colMeans()
posterior_means
hyperpars %>% gather() %>% ggplot(aes(x = value, fill=key)) + stat_density() + facet_wrap(~key)
```

```{r}
cov <- kernel
sigma.n <- posterior_means["obs_sd"]

cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% greta_array(obs$y)
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)

Ef <- calculate(Ef, 
                values = list(rbf_len = posterior_means["rbf_len"], 
                              rbf_var = posterior_means["rbf_var"])) %>% as.numeric()
Cf <- calculate(Cf, 
                values = list(rbf_len = posterior_means["rbf_len"], 
                              rbf_var = posterior_means["rbf_var"]))
```

```{r}
## Theory/Analytic
gp <- data.frame(x_predict = x_predict, Ef = Ef, sigma = 2*sqrt(diag(Cf)) )
## Draws

values <- MASS::mvrnorm(200, Ef, Cf)
dat <- data.frame(x=x_predict, t(values)) %>%  pivot_longer(-x)
```

```{r}
ggplot(dat,aes(x=x,y=value)) +
  geom_ribbon(data=gp, 
              aes(x=x_predict, y=Ef, ymin=Ef-sigma, ymax=Ef+sigma),
              fill="grey80") + # Var
  geom_line(aes(group=name), alpha=0.3, col=rgb(0.7, 0.1, 0.4, 0.1)) + #REPLICATES
  geom_line(data=gp,aes(x=x_predict,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
```

## Forecast from the estimated process

```{r}

## recompute
gp_forecast <- function(kernel, obs, x_t, steps){

  cov_xx_inv <- solve(cov(obs$x, obs$x))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)

}

## grid forecast: points in x_predict
gp <- data.frame(x = x_predict, Ef = Ef, sigma = 2*sqrt(diag(Cf)) )

## Compute decisions from the estimated process
